Allowing Stable Groups of Prosumers to Enter the Market through Proper Aggregation

### 2014-09-11 Article.tex
update

In a smart grid environement, we study coalitions of prosumers (agents that produce but also consume depending on time) which aim at selling energy to the grid. We reckon that the grid is able to specify two requirements (stability and minimum production) that enable a coalition to enter the market. We then design an algorithmm that seeks for uncorrelated profiles among the prosumers in order to form coalitions that are statistically stable enough to pass the grid requirements. The proposed algorithm uses complex system tools such as correlation graphs or clique percolation, and remains simple enough such that coalitions are formed in reasonable time. We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare, but that it is also more robust against grid minimum production threshold variations.

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}
% no \IEEEPARstart
One of the most basic ideas in the smart grid revolves around the introduction of communication means in the power grids, which could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}. Distributed energy resources (DER) such as wind turbines or photovoltaic panels are supposed to emerge and populate not only spread out farms, but also classical residential neighborhoods. Together with electric vehicles, and demand side management tools, they constitute the building blocks which will help turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}. Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.

It is rather clear that there should be communication flows between prosumers and the grid as most of the demand side management concepts suppose such a link \cite{WuHamedHuangBook2011}. Nevertheless as the expected number of active prosumers in the future seems quite large, the literature assume more and more communication between prosumers as this is a key ingredient to maintain more complex structures (such as coalitions), and thereby decrease the communication load on the grid’s side. Furthermore, it has been shown that coalitions in the smart grid provide nice ways of improving stability while posing exciting challenges in terms of network infrastructure and management. Self sufficient microgrids that can disconnect from the main grid \cite{Pahwa}, virtual power plants (VPP) that map small DER in sizeable and adjustable plants \cite{Braun} \cite{Ramchurn}, or coalitions of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn} are just a few examples of how coalitional concepts can enhance grid reliability.

In this paper, our focus is on how improving statistically the production stability of coalitions of prosumers while decreasing the communication needs for staying in a stable state. The idea is that, by using prediction techniques, coalitions can propose contract values to the grid, enabling it to schedule production on its side accordingly. Actually, contract values are particularly relevant in power exchanges contexts where energy is traded based on predictions (day-ahead estimations for instance). It seems obvious that, in such systems, participants should try to minimize prediction errors in order to maintain a stable state of grid operation, and that some penalty rules should be applied to ensure that coalitions are willing to report contract values correctly. Furthermore, as the coalition productions are supposed to come from (uncertain) renewables, the notion of reliability comes into play with a visible tradeoff between contract values worth and reliability.

In this paper, we seek at optimizing this tradeoff, that is forming coalitions able to announce “high contract values with high reliability”. We thus aim at:

%In this paper, our concern is on how prosumers should be aggregated in order to optimize this tradeoff, that is, how coalitions should be formed with the intention of announcing “high contract values with high reliability”. We thus aim at:

\begin{itemize}
\item Realistic representations (prosumers productions are non independent variables and come (indirectly) from real traces)
\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section \ref{sec:model})
\item Being (relatively) scalable with the number of prosumers
\end{itemize}

>>>>

In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid. It is paramount for the grid operation that the energy producers are able to sustain the grid requirements in term of stability and minimum production requirement in order to enter the energy market. We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have stable production level during the course of the bidding. We proposed an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalition that meet such complex constraints. We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations to due to changing weather condition for instance.

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
% Mantegna
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}
% no \IEEEPARstart
One of the key ideas in the smart grid revolves around the introduction of communication means inside the power grids. That could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}. Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge in only in solar farms, but also in residential areas. Together with electric vehicles, and demand side management tools, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}. Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G…) to ensure grid stability are called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.

There seems to have a clear consensus on the benefit of having a bidirectional communication flows between prosumers and the grid as most of the demand side management concepts have proposed \cite{WuHamedHuangBook2011}. Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication pattern (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management. The formation of coalition inside the smart grid environment could be applied to various agents such as the self sufficient micro-grids\cite{Pahwa}, the virtual power plants (VPP), sizable and adjustable plants \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}, are just a few use case where the coalition of multiple agents can enhance grid reliability and in the mean time reduce the communication burden.

In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probability of staying in acceptable range of production. By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly. Meeting contract values for the prosumer are especially relevant in power trade conditions, where energy is traded based on day-ahead predictions. In this context, participants should try to minimize its prediction error in order to maintain a stable state for the grid. They may even occur some penalty if the production deviate from its initial contract value. However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability in the context for renewable energy source will impact its net gain and for the power grid its help to maintain its reliability.

In this paper, we seek at optimizing this tradeoff, that is forming coalitions able to announce “high contract values with high reliability”. We thus aim at:
\begin{itemize}
\item Realistic prosumers productions with renewable production sources  from real traces base from weather measure.
\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section \ref{sec:model})
\item Being (relatively) scalable with the number of prosumers
\end{itemize}
